Comparing ChatGPT-4 to Physicians for Assigning Causes of Death from Verbal Autopsy: A Case Study of 6942 Deaths in Sierra Leone from 2019-2022

### 2024-02-07 manuscript/manuscript.tex
Add record distrib to appendices

Initially, 11,920 records from the HEAL-SL study \cite{njalauniversityHealthySierraLeone2023,carshon-marshChildMaternalAdult2022} were collected from dual-coded EVA, where each record was randomly coded by two different physicians that assigned CODs as International Classification of Diseases Revision 10 (ICD-10) codes \cite{worldhealthorganizationICD10InternationalStatistical2011}. Physicians were able to assign CODs for 11,820 of the 11,920 records, where 100 of these records could not be assigned a COD due to missing or inadequate information (e.g. low quality narrative, data loss). To determine if two codes were in physician agreement per record, codes were compared for similarity using Central Medical Evaluation Agreement 10 (CMEA-10) codes, which groups a range of similar ICD-10 codes together indicating codes in agreement \cite{aleksandrowiczPerformanceCriteriaVerbal2014} (see Additional File 1). When codes were not in agreement, a record enters the reconciliation phase, where the two physicians were provided reasoning and initial codes from each other to: (1) keep their initial code (2) assign the other physician's code or (3) assign a new code. If codes were not in agreement after the reconciliation phase, a record enters the adjudication phase, where a third senior physician evaluates both physicians' reasoning and codes before and after reconciliation, and assigns a final code based on their evaluation. The 11,820 physician coded records were further filtered for records where both physicians agreed on the assigned codes (records that were not reconciled or adjudicated) resulting in 6942 physician agreed records. Since computer models were compared to physicians in this study, there was more certainty that COD assignments agreed by both physicians were representative of physician assignment than when they disagreed \cite{barnettComparativeAccuracyDiagnosis2019,morrisFactorsAssociatedPhysician2010,hsiaoFactorsAssociatedPhysician2012}. The 6942 records were converted into Centre for Global Health Research 10 (CGHR-10) codes (see Additional File 2) that generalized ICD-10 codes into 19, 10, and 7 categories for the adult (12 years or older), child (28 days to 11 years), and neonatal (under 28 days) age groups.  After conversion, a final total of 6939 physician agreed records (3826 adult, 2636 child, and 477 neonatal) was used for modelling and performance evaluation, where three records were removed as their ICD-10 codes did not have a matching CGHR-10 code.

\subsubsection{Adult VA Records (12 Years or Older)}\label{methods-data-adult}

\subsubsection{Child VA Records (28 Days to 11 Years)}\label{methods-data-child}

\subsubsection{Neonatal VA Records (Under 28 Days)}\label{methods-data-neo}

@@

\begin{itemize}
    \item Additional file 1: Central Medical Evaluation Agreement 10 (CMEA-10) codes. ICD-10 code ranges considered in physician agreement. (.csv)
    \item Additional file 2: Centre for Global Health Research 10 (CGHR-10) codes. Codes grouping ICD-10 code ranges into generalized categories. (.csv)
\end{itemize}

@@

%\begin{appendices}

@@

%\end{appendices}

%%===========================================================================================%%
%% If you are submitting to one of the Nature Portfolio journals, using the eJP submission   %%
%% system, please include the references within the manuscript file itself. You may do this  %%
%% by copying the reference list from your .bbl file, paste it into the main manuscript .tex %%
%% file, and delete the associated \verb+\bibliography+ commands.                            %%
%%===========================================================================================%%
\bibliography{bibliography}% common bib file
%% if required, the content of .bbl file can be included here once bbl is generated
%%\input sn-article.bbl

>>>>

Initially, 11,920 records from the HEAL-SL study \cite{njalauniversityHealthySierraLeone2023,carshon-marshChildMaternalAdult2022} were collected from dual-coded EVA, where each record was randomly coded by two different physicians that assigned CODs as International Classification of Diseases Revision 10 (ICD-10) codes \cite{worldhealthorganizationICD10InternationalStatistical2011}. Physicians were able to assign CODs for 11,820 of the 11,920 records, where 100 of these records could not be assigned a COD due to missing or inadequate information (e.g. low quality narrative, data loss). To determine if two codes were in physician agreement per record, codes were compared for similarity using Central Medical Evaluation Agreement 10 (CMEA-10) codes, which groups a range of similar ICD-10 codes together indicating codes in agreement \cite{aleksandrowiczPerformanceCriteriaVerbal2014} (see Additional File 2). When codes were not in agreement, a record enters the reconciliation phase, where the two physicians were provided reasoning and initial codes from each other to: (1) keep their initial code (2) assign the other physician's code or (3) assign a new code. If codes were not in agreement after the reconciliation phase, a record enters the adjudication phase, where a third senior physician evaluates both physicians' reasoning and codes before and after reconciliation, and assigns a final code based on their evaluation. The 11,820 physician coded records were further filtered for records where both physicians agreed on the assigned codes (records that were not reconciled or adjudicated) resulting in 6942 physician agreed records. Since computer models were compared to physicians in this study, there was more certainty that COD assignments agreed by both physicians were representative of physician assignment than when they disagreed \cite{barnettComparativeAccuracyDiagnosis2019,morrisFactorsAssociatedPhysician2010,hsiaoFactorsAssociatedPhysician2012}. The 6942 records were converted into CGHR-10 codes (see Additional File 1) that generalized ICD-10 codes into 19, 10, and 7 categories for the adult (12 years or older), child (28 days to 11 years), and neonatal (under 28 days) age groups.  After conversion, a final total of 6939 physician agreed records (3826 adult, 2636 child, and 477 neonatal) was used for modelling and performance evaluation, where three records were removed as their ICD-10 codes did not have a matching CGHR-10 code.

@@

\begin{itemize}
    \item Additional file 1: Centre for Global Health Research 10 (CGHR-10) codes. Codes grouping ICD-10 code ranges into generalized categories. (.csv)
    \item Additional file 2: Central Medical Evaluation Agreement 10 (CMEA-10) codes. ICD-10 code ranges considered in physician agreement. (.csv)
\end{itemize}

@@

\newpage
\begin{appendices}

@@

\section{Physician Agreed Verbal Autopsy Records}\label{app-data}

The following figures visualize the distribution of all physician agreed records by age groups (Figure \ref{fig-data-agesex}) and 5-year age ranges (Figure \ref{fig-data-agerange}).

\begin{figure}[h]
\centering
\includegraphics[width=1\textwidth]{fig-data-agesex.pdf}
\caption{Number of physician agreed VA records by adult, child, and neonatal age groups.}\label{fig-data-agesex}
\end{figure}

\newpage
\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{fig-data-agerange.pdf}
\caption{Number of physician agreed VA records by 5-year age ranges.}\label{fig-data-agerange}
\end{figure}

\newpage
\section{Adult Physician Agreed Verbal Autopsy Records (12 Years or Older)}\label{app-data-adult}

The following figures visualize the distribution of adult physician agreed records by 5-year age ranges (Figure \ref{fig-data-agerange-adult}) and CGHR-10 cause of death code (Figure \ref{fig-data-cod-adult}).

\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{fig-data-agerange-adult.pdf}
\caption{Number of physician agreed adult (12 years or older) VA records by 5-year age ranges.}\label{fig-data-agerange-adult}
\end{figure}

\newpage
\begin{figure}[h]
\centering
\includegraphics[width=0.75\textwidth]{fig-data-cod-adult.pdf}
\caption{Number of physician agreed adult (12 years or older) VA records by CGHR-10 cause of death code.}\label{fig-data-cod-adult}
\end{figure}

\newpage
\section{Child Physician Agreed Verbal Autopsy Records (28 Days to 11 Years)}\label{app-data-child}

The following figures visualize the distribution of child physician agreed records by 5-year age ranges (Figure \ref{fig-data-agerange-child}) and CGHR-10 cause of death code (Figure \ref{fig-data-cod-child}).

\begin{figure}[h]
\centering
\includegraphics[width=0.75\textwidth]{fig-data-agerange-child.pdf}
\caption{Number of physician agreed child (28 days to 11 years) VA records by 5-year age ranges.}\label{fig-data-agerange-child}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{fig-data-cod-child.pdf}
\caption{Number of physician agreed child (28 days to 11 years) VA records by CGHR-10 cause of death code.}\label{fig-data-cod-child}
\end{figure}

\newpage
\section{Neonatal Physician Agreed Verbal Autopsy Records (Under 28 Days)}\label{app-data-neo}

The following figures visualize the distribution of neonatal physician agreed records by 5-year age ranges (Figure \ref{fig-data-agerange-neo}) and CGHR-10 cause of death code (Figure \ref{fig-data-cod-neo}).

\begin{figure}[h]
\centering
\includegraphics[width=0.8\textwidth]{fig-data-agerange-neo.pdf}
\caption{Number of physician agreed neonatal (less than 28 days) VA records by 5-year age ranges.}\label{fig-data-agerange-neo}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{fig-data-cod-neo.pdf}
\caption{Number of physician agreed neonatal (less than 28 days) VA records by CGHR-10 cause of death code.}\label{fig-data-cod-neo}
\end{figure}

\end{appendices}

%%===========================================================================================%%
%% If you are submitting to one of the Nature Portfolio journals, using the eJP submission   %%
%% system, please include the references within the manuscript file itself. You may do this  %%
%% by copying the reference list from your .bbl file, paste it into the main manuscript .tex %%
%% file, and delete the associated \verb+\bibliography+ commands.                            %%
%%===========================================================================================%%

\newpage
\bibliography{bibliography}% common bib file
%% if required, the content of .bbl file can be included here once bbl is generated
%%\input sn-article.bbl
